A Framework for Equivalence of Cast Calculi and a new Cast Representation

### 2020-09-27 paper.tex
label polarity renaming "car" to "fst" renaming "cdr" to "snd"

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
\usepackage{booktabs}   %% For formal tables:
%% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
%% http://ctan.org/pkg/subcaption
\usepackage{xcolor} %% For \color
\usepackage{todonotes} %% For \todo
\usepackage{mathtools} %% For \xRightarrow[sub]{sup}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stmaryrd} %% For \llbracket etc.
\usepackage{semantic}
\usepackage[scaled]{beramono}
\usepackage[T1]{fontenc}
\usepackage{natbib} %% For more citation commames (\cite etc.)
\usepackage{listings} %% For typesetting code
\newcommand{\lstsetgrift}{
	\lstset{%
		language=Lisp,%
		basicstyle=\ttfamily\small,%
		keywordstyle=\ttfamily\bfseries\small,%
		columns=flexible,%
		aboveskip=\smallskipamount,%
		belowskip=\smallskipamount,%
		frame=single,%
		xleftmargin=2pt,%
		escapeinside={/+}{+/},%
		mathescape=true,%
		keywords=[1]{define,if,:,Int,Bool,-,=,open},
		literate=%
		{->}{{$\rightarrow$}}2%
		{=>}{{$\Rightarrow$}}2%
		{lambda}{{$\lambda$}}1%
		{→}{{$\rightarrow$ }}2%
		{Γ}{{$\Gamma$ }}1%
		{⊢}{{$\vdash$ }}1%
	}%
}%%
\lstsetgrift

@@

% names
\newcommand{\LUD}{\ensuremath{\text{Lazy\;UD}}}
\newcommand{\LD}{\ensuremath{\text{Lazy\;D}}}
\newcommand{\CastADT}{\ensuremath{\text{Cast ADT}}}
\newcommand{\RMachine}[1]{\ensuremath{\mathcal{R}(#1)}}
\newcommand{\LDMachine}{\RMachine{\BLD}}
\newcommand{\LUDMachine}{\RMachine{\BLUD}}
\newcommand{\RBMachine}[0]{\RMachine{B}}
\newcommand{\SMachine}[1]{\ensuremath{\mathcal{S}(#1)}}

@@

%% expressions
\newcommand{\econ}[1]{#1}
\newcommand{\evar}[1]{#1}
\newcommand{\elam}[3]{\lambda#1:#2.\;#3}
\newcommand{\eapp}[2]{#1\;#2}
\newcommand{\etrue}[0]{\mathtt{true}}
\newcommand{\efalse}[0]{\mathtt{false}}
\newcommand{\eif}[3]{\mathtt{if}\;#1\;#2\;#3}
\newcommand{\econs}[2]{\mathtt{cons} \; #1 \; #2}
\newcommand{\ecar}[1]{\mathtt{car} \; #1}
\newcommand{\ecdr}[1]{\mathtt{cdr} \; #1}
%\newcommand{\ecast}[2]{#1 \langle #2 \rangle}
\newcommand{\ecast}[2]{\ensuremath{#1 : #2}}

@@

% frames
\newcommand{\closure}[2]{\langle#1,#2\rangle}
\newcommand{\FappXO}[2]{\eapp{\square}{\closure{#1}{#2}}}
\newcommand{\FappOX}[1]{\eapp{#1}{\square}}
\newcommand{\Fif}[3]{\eif{\square}{\closure{#1}{#3}}{\closure{#2}{#3}}}
\newcommand{\FconsXO}[2]{\econs{\square}{\closure{#1}{#2}}}
\newcommand{\FconsOX}[1]{\econs{#1}{\square}}
\newcommand{\Fcar}[0]{\ecar{\square}}
\newcommand{\Fcdr}[0]{\ecdr{\square}}
\newcommand{\Fcast}[1]{\ecast{\square}{#1}}

@@

% functions
\newcommand{\meet}[2]{\ensuremath{#1 \sqcap #2}}
\newcommand{\denote}[1]{\ensuremath{\llbracket #1 \rrbracket}}
\newcommand{\proxy}[2]{\ensuremath{\mathrm{proxy}(#1,#2)}}
\newcommand{\applyCast}[2]{\ensuremath{\mathrm{applyCast}(#1,#2)}}
\newcommand{\project}[4]{\ensuremath{\mathrm{project}(#1,#2,#3,#4)}}
\newcommand{\composegap}[3]{\ensuremath{#1 \fatsemi #2 \fatsemi #3}}
\newcommand{\mkhead}[2]{\ensuremath{\mathrm{mkProj}(#1,#2)}}
\newcommand{\mktail}[2]{\ensuremath{\mathrm{mkInj}(#1,#2)}}
\newcommand{\mkgap}[3]{\ensuremath{\mathrm{mkGap}(#1,#2,#3)}}
\newcommand{\compose}[2]{\ensuremath{#1 \fatsemi #2}}
\newcommand{\translate}[1]{\ensuremath{\lceil#1\rceil}}
\newcommand{\castToDyn}[2]{\ensuremath{\mathrm{castToDyn}(#1,#2)}}
\newcommand{\castFromDyn}[2]{\ensuremath{\mathrm{castFromDyn}(#1,#2)}}
\newcommand{\dom}[1]{\ensuremath{\mathrm{dom}(#1)}}
\newcommand{\cod}[1]{\ensuremath{\mathrm{cod}(#1)}}
\newcommand{\car}[1]{\ensuremath{\mathrm{fst}(#1)}}
\newcommand{\cdr}[1]{\ensuremath{\mathrm{snd}(#1)}}
\newcommand{\id}[1]{\ensuremath{\mathrm{id}(#1)}}
\newcommand{\idm}[1]{\ensuremath{\mathrm{id}_m(#1)}}
\newcommand{\idc}[1]{\ensuremath{\mathrm{id}_c(#1)}}
\newcommand{\mkCont}[1]{\ensuremath{\mathrm{mkK}(#1)}}
\newcommand{\extCont}[2]{\ensuremath{\mathrm{extK}(#1,#2)}}
% \newcommand{\applyCont}[2]{\ensuremath{\langle\langle#1,#2\rangle\rangle}}
\newcommand{\applyCont}[2]{\ensuremath{\mathrm{applyK}(#1,#2)}}

@@

In this section, we first review lazy cast calculi (Subsection~\ref{sec:cast-calculi}), where
we present an abstract machine that is employed in our framework. Then
we review coercions and their normal forms (Subsection~\ref{sec:coercions-in-normal-form}), which
motivates and inspires the design of hypercoercions.

@@

\begin{figure}
Syntax
\[
\begin{array}{lrcl}
\text{Blame Labels} & l \\
\text{Variables} & x,y,z \\
\stxrule{S,T}{Types}{
	\Tdyn \mid{}
	P
}
\stxrule{P,Q}{Pre-types}{
	\Pbool \mid
	\Pfunc{T}{T} \mid
	\Pprod{T}{T}
}
\stxrule{G,H}{Ground types}{
	\Gbool  \mid
	\Gfunc  \mid
	\Gprod
}
\stxrule{e}{Expressions}{
	\evar{x}        \mid
	\elam{x}{S}{e}  \mid
	\eapp{e}{e}     \mid
	\econs{e}{e}  \mid
	\ecar{e}      \mid
	\ecdr{e}
}
\stxrulecont{
	\ecast{e}{\ccast{S}{l}{T}}  \mid
	\etrue         \mid
	\efalse        \mid
	\eif{e}{e}{e}
}
\end{array}
\]

@@

The syntax and static semantics are the same for the \LD\ and
\LUD\ cast calculi. They are reviewed in \figref{fig:statics}.
As usual, the important features are the cast expressions,
$\ecast{e}{\ccast{S}{l}{T}}$, which are responsible for runtime type
checking. Polarities of blame labels is not treated but are straightforward to 
incorporate. The syntax and static semantics is the same as that of 
\citep{siek2009exploring} except for a few minor exceptions:

@@

\begin{figure}
\[
\begin{array}{lrcl}
\stxrule{v,u}{Values}{
  	\vdyn{I}{v}  \mid
    \vtrue       \mid
    \vfalse                     
}
\stxrulecont{
    \vfunc{x}{S}{e}{E}  \mid
	\vfuncc{v}{S}{T}{l}{S}{T}
}
\stxrulecont{
    \vcons{v}{v}               \mid
	\vconsc{v}{S}{T}{l}{S}{T}
}
\stxrule{I,J}{Injectable types}{
	\text{a subset of $P$ defined by the blame strategy $B$}
}
\stxrule{E}{Environments}{
	\text{a partial function}\;\{\langle x , v \rangle , \dots \}
}
\stxrule{F}{Frames}{
	\FappXO{e}{E}  \mid
	\FappOX{v}     \mid
	\Fcast{\ccast{S}{l}{T}} \mid
	\Fif{e}{e}{E}
}
\stxrulecont{
	\Fcar \mid
	\Fcdr \mid
	\FconsXO{e}{E}  \mid
	\FconsOX{v}
}
\stxrule{K}{Continuations}{
	\Knil         \mid
	\Kcons{F}{K}
}
\stxrule{s}{States}{
	\sexpr{e}{E}{K}  \mid
	\scont{v}{K}     \mid
	\shalt{v}        \mid
	\error{l}
}
\stxrule{o}{Observables}{
    \odyn    \mid
    \otrue   \mid
    \ofalse  \mid
    \ofunc   \mid
    \ocons   \mid
    \error{l}
}
\stxrule{r}{Cast results}{
    v \mid
    \error{l}
}
\end{array}
\]

@@

\begin{figure}
	\fbox{\reducetoR{B}{s}{s}} Transition
	\[
	\begin{array}{rclr}
	\figreduceto{\sexpr{\evar{x}}{E}{K}}{\scont{E(x)}{K}}{}
	\figreduceto{\sexpr{\elam{x}{S}{e}}{E}{K}}{\scont{\vfunc{x}{S}{e}{E}}{K}}{}
	\figreduceto{\sexpr{\eapp{e_1}{e_2}}{E}{K}}{
		\sexpr{e_1}{E}{\Kcons{\FappXO{e_2}{E}}{K}}
	}{}
	\figreduceto{\sexpr{\etrue}{E}{K}}{\scont{\vtrue}{K}}{}
	\figreduceto{\sexpr{\efalse}{E}{K}}{\scont{\vfalse}{K}}{}
	\figreduceto{\sexpr{\eif{e_1}{e_2}{e_3}}{E}{K}}{
		\scont{e_1}{\Kcons{\Fif{e_2}{e_3}{E}}{K}}
	}{}
	\figreduceto{\sexpr{\econs{e_1}{e_2}}{E}{K}}{
		\sexpr{e_1}{E}{\Kcons{\FconsXO{e_2}{E}}{K}}
	}{}
	\figreduceto{\sexpr{\ecar{e}}{E}{K}}{\sexpr{e}{E}{\Kcons{\Fcar}{K}}}{}
	\figreduceto{\sexpr{\ecdr{e}}{E}{K}}{\sexpr{e}{E}{\Kcons{\Fcdr}{K}}}{}
	\figreducetoH{
		\sexpr{\ecast{e}{\ccast{S}{l}{T}}}{E}{K}
	}{
		\sexpr{e}{E}{\Kcons{\Fcast{\ccast{S}{l}{T}}}{K}}
	}{}
	\figreducetoH{\scont{v}{\Kcons{\Fcast{\ccast{S}{l}{T}}}{K}}}{
		\applyCast{v}{\ccast{S}{l}{T}} \; \mbind \; \lambda u. \scont{u}{K}
	}{}
	\figreduceto{\scont{v_1}{\Kcons{\FappXO{e_2}{E}}{K}}}{\sexpr{e_2}{E}{
			\Kcons{\FappOX{v_1}}{K}}
	}{}
	\figreduceto{\scont{v}{\Kcons{\FappOX{\vfunc{x}{S}{e}{E}}}{K}}}{
		\sexpr{e}{E[x:=v]}{K}
	}{}
	\figreducetoH{\scont{v}{\Kcons{\FappOX{(\vfuncc{u}{S_1}{T_1}{l}{S_2}{T_2})}}{K}}}{
		\scont{v}{
			\Kcons{\Fcast{\ccast{S_2}{l}{S_1}}}{
				\Kcons{\FappOX{u}}{
					\Kcons{\Fcast{\ccast{T_1}{l}{T_2}}}{K}}}}
	}{}
	\figreduceto{\scont{\vtrue}{\Kcons{\Fif{e_2}{e_3}{E}}{K}}}{\sexpr{e_2}{E}{K}}{}
	\figreduceto{\scont{\vfalse}{\Kcons{\Fif{e_2}{e_3}{E}}{K}}}{\sexpr{e_3}{E}{K}}{}
	\figreduceto{\scont{v_1}{\Kcons{\FconsXO{e_2}{E}}{K}}}{
		\sexpr{e_2}{E}{\Kcons{\FconsOX{v_1}}{K}}
	}{}
	\figreduceto{\scont{v_2}{\Kcons{\FconsOX{v_1}}{K}}}{
		\scont{\vcons{v_1}{v_2}}{K}
	}{}
	\figreduceto{\scont{\vcons{v_1}{v_2}}{\Kcons{\Fcar}{K}}}{\scont{v_1}{K}}{}
	\figreducetoH{\scont{\vconsc{u}{S_1}{T_1}{l}{S_2}{T_2}}{\Kcons{\Fcar}{K}}}{
		\scont{u}{
			\Kcons{\Fcar}{
				\Kcons{\Fcast{\ccast{S_1}{l}{S_2}}}{K}}}	
	}{}
	\figreduceto{\scont{\vcons{v_1}{v_2}}{\Kcons{\Fcdr}{K}}}{\scont{v_2}{K}}{}
	\figreducetoH{\scont{\vconsc{u}{S_1}{T_1}{l}{S_2}{T_2}}{\Kcons{\Fcdr}{K}}}{
		\scont{u}{
			\Kcons{\Fcdr}{
				\Kcons{\Fcast{\ccast{T_1}{l}{T_2}}}{K}}}	
	}{}
	\figreduceto{\scont{v}{\Knil}}{\shalt{v}}{}
	\end{array}
	\]

@@

\subsection{Coercions and Normal Forms} 
\label{sec:coercion-calculus}

@@

\begin{figure}
	\[
	\begin{array}{lrcl}
	\stxrule{I,J}{Injectable types (\LD)}{
		\Pfunc{T}{T} \mid \Pbool}
	\stxrule{I,J}{Injectable types (\LUD)}{
		\Pfunc{\Tdyn}{\Tdyn} \mid \Pbool
	}
	\stxrule{c,d}{Coercions}{
		\cnfid{T}       \mid
		\cnffunc{c}{d}  \mid
		\cnfprod{c}{d}  \mid
		\cnffail{l}     \mid
		\cnfinj{I}      \mid
		\cnfproj{I}{l}  \mid
		\cnfseq{c}{d}
	}
	\stxrule{s,t}{Coercions in normal form}{
		\cnfid{\Tdyn}       \mid
		\cnfseq{\cnfproj{I}{l}}{i} \mid
		i
	}
	\stxrule{i}{Final Coercions}{
		\cnffail{l} \mid
		\cnfseq{g}{\cnfinj{I}} \mid
		g
	}
	\stxrule{g}{Middle Coercions}{
		\cnfid{\Pbool} \mid
		\cnffunc{s}{t} \mid
		\cnfprod{s}{t}
	}
	\end{array}
	\]
	\caption{Syntax of coercions and normal forms {\`a} la 
		\citep{siek2012interpretations}.}
	\label{fig:normal-coercion}
\end{figure}

\figref{fig:normal-coercion} reviews the grammar for coercions,
written $c$.
%
To review, an injection $\cnfinj{I}$ takes a value from an injectable type
$I$ to type $\Tdyn$. An injectable type is simply a type that can be
cast directly to and from $\Tdyn$. The
definition of injectable types depends on the blame strategy, as in 
Chapter~\ref{chap:background}.
%
A projection $I?^l$ takes a value from type $\Tdyn$ to type $I$, or
halts the program and blames $l$ if the value is inconsistent with 
the target type.
The coercion \cnfid{T} is the identity, $\bot^{l}$ is the coercion that
always fails and blames $l$, and $(\cnfseq{c}{d})$ applies $c$ and
then $d$ in sequence. The function coercion $\cnffunc{c}{d}$
applies $c$ to the argument of a function and $d$ to the return
value.

@@

@@

\fbox{$\car{g}=g$}
\[
\begin{array}{rclr}
\figequalto{\car{\gnone}}{\gnone}{}
\figequalto{\car{\gsome{\Pprod{S_1}{T_1}}{l}{\Pprod{S_2}{T_2}}}}{
	\gsome{S_1}{l}{S_2}
}{}
\end{array}
\]

>>>>

% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
\usepackage{booktabs}   %% For formal tables:
%% http://ctan.org/pkg/booktabs
%\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
%% http://ctan.org/pkg/subcaption
\usepackage{xcolor} %% For \color
\usepackage{todonotes} %% For \todo
\usepackage{mathtools} %% For \xRightarrow[sub]{sup}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stmaryrd} %% For \llbracket etc.
\usepackage{semantic}
\usepackage[scaled]{beramono}
\usepackage[T1]{fontenc}
\usepackage[numbers]{natbib} %% For more citation commames (\cite etc.)
\usepackage{listings} %% For typesetting code
\newcommand{\lstsetgrift}{
	\lstset{%
		language=Lisp,%
		basicstyle=\ttfamily\small,%
		keywordstyle=\ttfamily\bfseries\small,%
		columns=flexible,%
		aboveskip=\smallskipamount,%
		belowskip=\smallskipamount,%
		frame=single,%
		xleftmargin=2pt,%
		escapeinside={/+}{+/},%
		mathescape=true,%
		keywords=[1]{define,if,:,Int,Bool,-,=,open},
		literate=%
		{->}{{$\rightarrow$}}2%
		{=>}{{$\Rightarrow$}}2%
		{lambda}{{$\lambda$}}1%
		{→}{{$\rightarrow$ }}2%
		{Γ}{{$\Gamma$ }}1%
		{⊢}{{$\vdash$ }}1%
	}%
}%%
\lstsetgrift

@@

% names
\newcommand{\LUD}{\ensuremath{\text{Lazy\;UD}}}
\newcommand{\LD}{\ensuremath{\text{Lazy\;D}}}
\newcommand{\CastADT}{\ensuremath{\text{Cast ADT}}}
\newcommand{\RMachine}[1]{\ensuremath{\mathcal{R}(#1)}}
\newcommand{\LDMachine}{\RMachine{\BLD}}
\newcommand{\LUDMachine}{\RMachine{\BLUD}}
\newcommand{\RBMachine}[0]{\RMachine{B}}
\newcommand{\SMachine}[1]{\ensuremath{\mathcal{S}(#1)}}
\newcommand{\effCEK}[1]{\ensuremath{\mathcal{S}(#1)}}

@@

%% expressions
\newcommand{\econ}[1]{#1}
\newcommand{\evar}[1]{#1}
\newcommand{\elam}[3]{\lambda#1:#2.\;#3}
\newcommand{\eapp}[2]{#1\;#2}
\newcommand{\etrue}[0]{\mathtt{true}}
\newcommand{\efalse}[0]{\mathtt{false}}
\newcommand{\eif}[3]{\mathtt{if}\;#1\;#2\;#3}
\newcommand{\econs}[2]{\mathtt{cons} \; #1 \; #2}
\newcommand{\efst}[1]{\mathtt{fst} \; #1}
\newcommand{\esnd}[1]{\mathtt{snd} \; #1}
%\newcommand{\ecast}[2]{#1 \langle #2 \rangle}
\newcommand{\ecast}[2]{\ensuremath{#1 : #2}}

@@

% frames
\newcommand{\closure}[2]{\langle#1,#2\rangle}
\newcommand{\FappXO}[2]{\eapp{\square}{\closure{#1}{#2}}}
\newcommand{\FappOX}[1]{\eapp{#1}{\square}}
\newcommand{\Fif}[3]{\eif{\square}{\closure{#1}{#3}}{\closure{#2}{#3}}}
\newcommand{\FconsXO}[2]{\econs{\square}{\closure{#1}{#2}}}
\newcommand{\FconsOX}[1]{\econs{#1}{\square}}
\newcommand{\Ffst}[0]{\efst{\square}}
\newcommand{\Fsnd}[0]{\esnd{\square}}
\newcommand{\Fcast}[1]{\ecast{\square}{#1}}

@@

% functions
\newcommand{\meet}[2]{\ensuremath{#1 \sqcap #2}}
\newcommand{\denote}[1]{\ensuremath{\llbracket #1 \rrbracket}}
\newcommand{\proxy}[2]{\ensuremath{\mathrm{proxy}(#1,#2)}}
\newcommand{\applyCast}[2]{\ensuremath{\mathrm{applyCast}(#1,#2)}}
\newcommand{\project}[4]{\ensuremath{\mathrm{project}(#1,#2,#3,#4)}}
\newcommand{\composegap}[3]{\ensuremath{#1 \fatsemi #2 \fatsemi #3}}
\newcommand{\mkhead}[2]{\ensuremath{\mathrm{mkProj}(#1,#2)}}
\newcommand{\mktail}[2]{\ensuremath{\mathrm{mkInj}(#1,#2)}}
\newcommand{\mkgap}[3]{\ensuremath{\mathrm{mkGap}(#1,#2,#3)}}
\newcommand{\compose}[2]{\ensuremath{#1 \fatsemi #2}}
\newcommand{\translate}[1]{\ensuremath{\lceil#1\rceil}}
\newcommand{\castToDyn}[2]{\ensuremath{\mathrm{castToDyn}(#1,#2)}}
\newcommand{\castFromDyn}[2]{\ensuremath{\mathrm{castFromDyn}(#1,#2)}}
\newcommand{\dom}[1]{\ensuremath{\mathrm{dom}(#1)}}
\newcommand{\cod}[1]{\ensuremath{\mathrm{cod}(#1)}}
\newcommand{\fst}[1]{\ensuremath{\mathrm{fst}(#1)}}
\newcommand{\snd}[1]{\ensuremath{\mathrm{snd}(#1)}}
\newcommand{\id}[1]{\ensuremath{\mathrm{id}(#1)}}
\newcommand{\idm}[1]{\ensuremath{\mathrm{id}_m(#1)}}
\newcommand{\idc}[1]{\ensuremath{\mathrm{id}_c(#1)}}
\newcommand{\mkCont}[1]{\ensuremath{\mathrm{mkK}(#1)}}
\newcommand{\extCont}[2]{\ensuremath{\mathrm{extK}(#1,#2)}}
% \newcommand{\applyCont}[2]{\ensuremath{\langle\langle#1,#2\rangle\rangle}}
\newcommand{\applyCont}[2]{\ensuremath{\mathrm{applyK}(#1,#2)}}

@@

In this section, we first review lazy cast calculi (Subsection~\ref{sec:cast-calculi}), where
we present an abstract machine that is employed in our framework. Then
we review coercions and their normal forms (Subsection~\ref{sec:coercions}), which
motivates and inspires the design of hypercoercions.

@@

\begin{figure}
Syntax
\[
\begin{array}{lrcl}
\text{Blame Labels} & l,\bar{l} \\
\text{Variables} & x,y,z \\
\stxrule{S,T}{Types}{
	\Tdyn \mid{}
	P
}
\stxrule{P,Q}{Pre-types}{
	\Pbool \mid
	\Pfunc{T}{T} \mid
	\Pprod{T}{T}
}
\stxrule{G,H}{Ground types}{
	\Gbool  \mid
	\Gfunc  \mid
	\Gprod
}
\stxrule{e}{Expressions}{
	\evar{x}        \mid
	\elam{x}{S}{e}  \mid
	\eapp{e}{e}     \mid
	\econs{e}{e}  \mid
	\efst{e}      \mid
	\esnd{e}
}
\stxrulecont{
	\ecast{e}{\ccast{S}{l}{T}}  \mid
	\etrue         \mid
	\efalse        \mid
	\eif{e}{e}{e}
}
\end{array}
\]

@@

The syntax and static semantics are the same for the \LD\ and
\LUD\ cast calculi. They are reviewed in \figref{fig:statics}.
As usual, the important features are the cast expressions,
$\ecast{e}{\ccast{S}{l}{T}}$, which are responsible for runtime type
checking. Blame labels can be positive (blaming the expression inside the cast)
or negative (blaming the context outside the cast), as in \cite{Wadler:2009qv,Siek:2015ab}.
The syntax and static semantics is the same as that of 
\citep{siek2009exploring} except for a few minor exceptions:

@@

\begin{figure}
\[
\begin{array}{lrcl}
\stxrule{v,u}{Values}{
  	\vdyn{I}{v}  \mid
    \vtrue       \mid
    \vfalse                     
}
\stxrulecont{
    \vfunc{x}{S}{e}{E}  \mid
	\vfuncc{v}{S}{T}{l}{S}{T}
}
\stxrulecont{
    \vcons{v}{v}               \mid
	\vconsc{v}{S}{T}{l}{S}{T}
}
\stxrule{I,J}{Injectable types}{
	\text{a subset of $P$ defined by the blame strategy $B$}
}
\stxrule{E}{Environments}{
	\text{a partial function}\;\{\langle x , v \rangle , \dots \}
}
\stxrule{F}{Frames}{
	\FappXO{e}{E}  \mid
	\FappOX{v}     \mid
	\Fcast{\ccast{S}{l}{T}} \mid
	\Fif{e}{e}{E}
}
\stxrulecont{
	\Ffst \mid
	\Fsnd \mid
	\FconsXO{e}{E}  \mid
	\FconsOX{v}
}
\stxrule{K}{Continuations}{
	\Knil         \mid
	\Kcons{F}{K}
}
\stxrule{s}{States}{
	\sexpr{e}{E}{K}  \mid
	\scont{v}{K}     \mid
	\shalt{v}        \mid
	\error{l}
}
\stxrule{o}{Observables}{
    \odyn    \mid
    \otrue   \mid
    \ofalse  \mid
    \ofunc   \mid
    \ocons   \mid
    \error{l}
}
\stxrule{r}{Cast results}{
    v \mid
    \error{l}
}
\end{array}
\]

@@

\begin{figure}
	\fbox{\reducetoR{B}{s}{s}} Transition
	\[
	\begin{array}{rclr}
	\figreduceto{\sexpr{\evar{x}}{E}{K}}{\scont{E(x)}{K}}{}
	\figreduceto{\sexpr{\elam{x}{S}{e}}{E}{K}}{\scont{\vfunc{x}{S}{e}{E}}{K}}{}
	\figreduceto{\sexpr{\eapp{e_1}{e_2}}{E}{K}}{
		\sexpr{e_1}{E}{\Kcons{\FappXO{e_2}{E}}{K}}
	}{}
	\figreduceto{\sexpr{\etrue}{E}{K}}{\scont{\vtrue}{K}}{}
	\figreduceto{\sexpr{\efalse}{E}{K}}{\scont{\vfalse}{K}}{}
	\figreduceto{\sexpr{\eif{e_1}{e_2}{e_3}}{E}{K}}{
		\scont{e_1}{\Kcons{\Fif{e_2}{e_3}{E}}{K}}
	}{}
	\figreduceto{\sexpr{\econs{e_1}{e_2}}{E}{K}}{
		\sexpr{e_1}{E}{\Kcons{\FconsXO{e_2}{E}}{K}}
	}{}
	\figreduceto{\sexpr{\efst{e}}{E}{K}}{\sexpr{e}{E}{\Kcons{\Ffst}{K}}}{}
	\figreduceto{\sexpr{\esnd{e}}{E}{K}}{\sexpr{e}{E}{\Kcons{\Fsnd}{K}}}{}
	\figreducetoH{
		\sexpr{\ecast{e}{\ccast{S}{l}{T}}}{E}{K}
	}{
		\sexpr{e}{E}{\Kcons{\Fcast{\ccast{S}{l}{T}}}{K}}
	}{}
	\figreducetoH{\scont{v}{\Kcons{\Fcast{\ccast{S}{l}{T}}}{K}}}{
		\applyCast{v}{\ccast{S}{l}{T}} \; \mbind \; \lambda u. \scont{u}{K}
	}{}
	\figreduceto{\scont{v_1}{\Kcons{\FappXO{e_2}{E}}{K}}}{\sexpr{e_2}{E}{
			\Kcons{\FappOX{v_1}}{K}}
	}{}
	\figreduceto{\scont{v}{\Kcons{\FappOX{\vfunc{x}{S}{e}{E}}}{K}}}{
		\sexpr{e}{E[x:=v]}{K}
	}{}
	\figreducetoH{\scont{v}{\Kcons{\FappOX{(\vfuncc{u}{S_1}{T_1}{l}{S_2}{T_2})}}{K}}}{
		\scont{v}{
			\Kcons{\Fcast{\ccast{S_2}{\bar{l}}{S_1}}}{
				\Kcons{\FappOX{u}}{
					\Kcons{\Fcast{\ccast{T_1}{l}{T_2}}}{K}}}}
	}{}
	\figreduceto{\scont{\vtrue}{\Kcons{\Fif{e_2}{e_3}{E}}{K}}}{\sexpr{e_2}{E}{K}}{}
	\figreduceto{\scont{\vfalse}{\Kcons{\Fif{e_2}{e_3}{E}}{K}}}{\sexpr{e_3}{E}{K}}{}
	\figreduceto{\scont{v_1}{\Kcons{\FconsXO{e_2}{E}}{K}}}{
		\sexpr{e_2}{E}{\Kcons{\FconsOX{v_1}}{K}}
	}{}
	\figreduceto{\scont{v_2}{\Kcons{\FconsOX{v_1}}{K}}}{
		\scont{\vcons{v_1}{v_2}}{K}
	}{}
	\figreduceto{\scont{\vcons{v_1}{v_2}}{\Kcons{\Ffst}{K}}}{\scont{v_1}{K}}{}
	\figreducetoH{\scont{\vconsc{u}{S_1}{T_1}{l}{S_2}{T_2}}{\Kcons{\Ffst}{K}}}{
		\scont{u}{
			\Kcons{\Ffst}{
				\Kcons{\Fcast{\ccast{S_1}{l}{S_2}}}{K}}}	
	}{}
	\figreduceto{\scont{\vcons{v_1}{v_2}}{\Kcons{\Fsnd}{K}}}{\scont{v_2}{K}}{}
	\figreducetoH{\scont{\vconsc{u}{S_1}{T_1}{l}{S_2}{T_2}}{\Kcons{\Fsnd}{K}}}{
		\scont{u}{
			\Kcons{\Fsnd}{
				\Kcons{\Fcast{\ccast{T_1}{l}{T_2}}}{K}}}	
	}{}
	\figreduceto{\scont{v}{\Knil}}{\shalt{v}}{}
	\end{array}
	\]

@@

\subsection{Coercions and Normal Forms} 
\label{sec:coercions}

@@

\begin{figure}
	\[
	\begin{array}{lrcl}
	\stxrule{I,J}{Injectable types (\LD)}{P}
	\stxrule{I,J}{Injectable types (\LUD)}{G}
	\stxrule{c,d}{Coercions}{
		\cnfid{T}       \mid
		\cnffunc{c}{d}  \mid
		\cnfprod{c}{d}  \mid
		\cnffail{l}     \mid
		\cnfinj{I}      \mid
		\cnfproj{I}{l}  \mid
		\cnfseq{c}{d}
	}
	\stxrule{s,t}{Coercions in normal form}{
		\cnfid{\Tdyn}       \mid
		\cnfseq{\cnfproj{I}{l}}{i} \mid
		i
	}
	\stxrule{i}{Final Coercions}{
		\cnffail{l} \mid
		\cnfseq{g}{\cnfinj{I}} \mid
		g
	}
	\stxrule{g}{Middle Coercions}{
		\cnfid{\Pbool} \mid
		\cnffunc{s}{t} \mid
		\cnfprod{s}{t}
	}
	\end{array}
	\]
	\caption{Syntax of coercions and normal forms {\`a} la 
		\citep{siek2012interpretations}.}
	\label{fig:normal-coercion}
\end{figure}

\figref{fig:normal-coercion} reviews the grammar for coercions,
written $c$.
%
To review, an injection $\cnfinj{I}$ takes a value from an injectable type
$I$ to type $\Tdyn$. An injectable type is simply a type that can be
cast directly to and from $\Tdyn$. The
definition of injectable types depends on the blame strategy, as in 
Subsection~\ref{sec:cast-calculi}.
%
A projection $I?^l$ takes a value from type $\Tdyn$ to type $I$, or
halts the program and blames $l$ if the value is inconsistent with 
the target type.
The coercion \cnfid{T} is the identity, $\bot^{l}$ is the coercion that
always fails and blames $l$, and $(\cnfseq{c}{d})$ applies $c$ and
then $d$ in sequence. The function coercion $\cnffunc{c}{d}$
applies $c$ to the argument of a function and $d$ to the return
value.

@@

\fbox{$\bisimv{v}{v}{T}$}
\begin{gather*}
\infr{
	\expressiontyping{\Gamma,x : S_1}{e}{T_1}
	\qquad
	\bisimE{\Gamma}{E}{E'}
	\\
	\bisimc{
		(\ccast{S_n}{\bar{l}_{n-1}}{S_{n-1}} \ldots
		\ccast{S_2}{\bar{l}_1}{S_1})
	}{c}{S_n}{S_1}
	\\
	\bisimc{
		(		\ccast{T_1}{l_{1}}{T_2} \ldots
		\ccast{T_{n-1}}{l_{n-1}}{T_n}
		)	}{d}{T_1}{T_n}
	\\
	v = \vfuncc{
		\vfuncc{
			\vfunc{x}{S_1}{e}{E}
		}{S_1}{T_1}{l_1}{S_2}{T_2}
		\ldots
	}{S_{n-1}}{T_{n-1}}{l_{n-1}}{S_n}{T_n}
}{
	\bisimv{v}{
		(\ecast{\vfunc{x}{S_1}{e}{E'}}{\Pfunc{c}{d}})
	}{
		\Pfunc{S_n}{T_n}}
}
\\[1ex]
\infr{
	\bisimv{v_1}{v_1'}{S_1} \qquad
	\bisimv{v_2}{v_2'}{T_1} \\
	\bisimc{
		(\ccast{S_1}{l_{1}}{S_2} \ldots
		\ccast{S_{n-1}}{l_{n-1}}{S_n})
	}{c}{S_1}{S_n} \\
	\bisimc{
		(\ccast{T_1}{l_{1}}{T_2} \ldots
		\ccast{T_{n-1}}{l_{n-1}}{T_n})
	}{d}{T_1}{T_n}
	\\ v = \vconsc{
		\vconsc{
			\vcons{v_1}{u_1}
		}{S_1}{T_1}{l_1}{S_2}{T_2}
		\ldots
	}{S_{n-1}}{T_{n-1}}{l_{n-1}}{S_n}{T_n}
}{
	\bisimv{
		v
	}{
		(\ecast{\vcons{v_2}{u_2}}{\Pprod{c}{d}})
	}{
		\Pprod{S_n}{T_n}}
}
\end{gather*}

@@

\fbox{$\fst{g}=g$}
\[
\begin{array}{rclr}
\figequalto{\fst{\gnone}}{\gnone}{}
\figequalto{\fst{\gsome{\Pprod{S_1}{T_1}}{l}{\Pprod{S_2}{T_2}}}}{
	\gsome{S_1}{l}{S_2}
}{}
\end{array}
\]
